---
title: "Harbor seals analysis"
author: "Kyra Bankhead"
date: "2022-12-14"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this markdown I will:

1. Find the best distribution to use for GLMs.

2. Run GLMS and AICs to find the most appropriate model and predictors.

3. Create visualization graphs for each site.

## Check for Appropriate Distribution

```{r density, echo=FALSE, message=FALSE, warning=FALSE}
full.data<-read.csv("../data/full.data.csv")
# Load packages
require(ggplot2)

# Check distribution
ggplot(full.data, aes(x=seals)) + 
  geom_histogram(aes(y=..density..), colour="black", fill="white")+
  geom_density()+stat_density(alpha=.2,adjust = 1, fill="#FF6666")+xlab("Number of Seals Hauled-out")+ylab("Density")+theme(panel.background = element_blank())

```

Negative binomial or poisson would be the best fit. Let's check if the mean is equal to the variance and if there is zero inflation.

```{r choice, echo=TRUE}
full.data<-read.csv("../data/full.data.csv")
require(performance)

var(full.data$seals)
mean(full.data$seals)

# Make sure overdispersion is detected with full model
mod<- glm(seals ~ site*noise + month + tide + time, data = full.data, family = "poisson")
check_overdispersion(mod) 
# Check for zero inflation
check_zeroinflation(mod)
```

The variance is way higher than the mean which was confirmed by the Pearson chi-squared test. Additionally there was zero-inflation present in the model. Therefore the Poisson distribution is not applicable here and we will continue with the negative binomial distribution.

## Create GLMs and find best model with AICc

To test whether noise affects the number of seals hauled-out by site, I will insert an interaction between noise level and site.

```{r glm, echo=FALSE, message=FALSE, warning=FALSE}
full.data<-read.csv("../data/full.data.csv")
# Load packages
require(MASS) #for glm.nb()
require(AICcmodavg) #for AICc

model1<- glm.nb(seals ~ 1, data = full.data)
model2<- glm.nb(seals ~ site*noise + month + tide + time, data = full.data)
model3<- glm.nb(seals ~ site*noise + month + time, data = full.data)
model4<- glm.nb(seals ~ site*noise + month, data = full.data)
# Create list
models<-list(model1, model2, model3, model4)

## Calculate AICc with glm of models
Modnames<- c("seals ~ 1",
             "seals ~ site*noise + month + tide + time",
             "seals ~ site*noise + month + time",
             "seals ~ site*noise + month")

aictab(models, modnames = Modnames,
                  second.ord = TRUE, 
                  nobs = NULL, sort = TRUE, c.hat = 1) # Looks like site*noise + month + time are the best predictors


```

Looks like the best model will contain month, noise, site and time as predictors. This is the summary of that model:

```{r model, echo=FALSE}
full.data<-read.csv("../data/full.data.csv")
# Load packages
require(MASS) #for glm.nb()

model3<- glm.nb(seals ~ site*noise + month + time, data = full.data)
summary(model3) 
```

* Month and time are significant predictors for how many harbor seals haul-out.

* Site and noise are significant predictors for the number of harbor seals hauled-out. The effect of noise on the number of seals haul-out depends on what site they are located in.

```{r interact, echo=FALSE, message=FALSE, warning=FALSE}
full.data<-read.csv("../data/full.data.csv")

# What does the interaction between site and noise look like?
interaction.plot(x.factor = full.data$noise, #x-axis variable
                 trace.factor = full.data$site, #variable for lines
                 response = full.data$seals, #y-axis variable
                 fun = median, #metric to plot
                 ylab = "Number of Seals Hauled-out",
                 xlab = "Noise Level (dB)",
                 col = c("pink", "blue"),
                 lty = 1, #line type
                 lwd = 2, #line width
                 trace.label = "Site")
```

## Use model output to predict response variable

```{r predict, echo=FALSE}
full.data<-read.csv("../data/full.data.csv")

newdata <- data.frame(noise = mean(full.data$noise), 
                       site = factor(1:2, levels = 1:2, 
                                      labels = unique(full.data$site)))
m3<- glm.nb(seals ~ site * noise, data = full.data)
newdata$phat <- predict.glm(m3, newdata, type = "response")
newdata
```

In the output above, we see that the predicted number of seals hauled-out for the Marina is about 17.03, holding noise level at its mean. The predicted number of events for the Waterfront is lower at 8.60.

Below we will obtain the mean predicted number of seals hauled-out for values of noise across its entire range for each site and graph these.

```{r range, echo=FALSE, message=FALSE, warning=FALSE}
full.data<-read.csv("../data/full.data.csv")
require(ggplot2)

# Use model to predict the response variable 
newdata <- data.frame(noise = mean(full.data$noise), 
                       site = factor(1:2, levels = 1:2, 
                                      labels = unique(full.data$site)))
m3<- glm.nb(seals ~ site * noise, data = full.data)
newdata$phat <- predict.glm(m3, newdata, type = "response")
newdata

# Plot 
newdata2 <- data.frame(
  noise = rep(seq(from = min(full.data$noise), to = max(full.data$noise), length.out = 100), 2),
  site = factor(rep(1:2, each = 100), levels = 1:2, labels =
                  unique(full.data$site)))

newdata2 <- cbind(newdata2, predict(m3, newdata2, type = "link", se.fit=TRUE))
newdata2 <- within(newdata2, {
  seals <- exp(fit)
  LL <- exp(fit - 1.96 * se.fit)
  UL <- exp(fit + 1.96 * se.fit)
})

ggplot(newdata2, aes(noise, seals)) +
  geom_ribbon(aes(ymin = LL, ymax = UL, fill = site), alpha = .25) +
  geom_line(aes(colour = site), linewidth = 2) +
  labs(x = "Noise Level (dB)", y = "Predicted Number of Seals Hauled-out")

```

The graph shows the log linear model of the expected seals counts across the range of noise levels, for each site along with 95 percent confidence intervals. Note that what is plotted are the expected values, not the log of the expected values.
