---
title: "Harbor seals analysis"
author: "Kyra Bankhead"
date: "2022-12-14"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Waterfront-Marina GLM analysis

In this markdown I will:

1. Find the best distribution to use for GLMs.

2. Run GLMS and AICs to find the most appropriate model and predictors.

3. Create visualization graphs for each site.

### Check Histogram

```{r density, echo=FALSE, message=FALSE}
setwd("C:/Users/bankh/My_Repos/habor-seal/data")
full.data<-read.csv("full.data.csv")
# Load packages
require(ggplot2)

# Check distribution
ggplot(full.data, aes(x=seals)) + 
  geom_histogram(aes(y=..density..), colour="black", fill="white")+
  geom_density()+stat_density(alpha=.2,adjust = 1, fill="#FF6666")+xlab("Number of Seals Hauled-out")+ylab("Density")+theme(panel.background = element_blank())

```

A negative binomial model would fit this data best.

## Create GLMs and find best model with AICc
To test whether noise affects the number of seals hauled-out by site, I will insert an interaction between noise level and site.

```{r glm, echo=FALSE}
setwd("C:/Users/bankh/My_Repos/habor-seal/data")
full.data<-read.csv("full.data.csv")
# Load packages
require(MASS) #for glm.nb()

model1<- glm.nb(seals ~ 1, data = full.data)
model2<- glm.nb(seals ~ site*noise + month + tide + time, data = full.data)
model3<- glm.nb(seals ~ site*noise + month + time, data = full.data)
model4<- glm.nb(seals ~ site*noise + month, data = full.data)
# Create list
models<-list(model1, model2, model3, model4)

## Find the AICc function
source("../code/Functions.R")

## Calculate AICc with glm of models
AICc(models) # Looks like site*noise + month + time are the best predictors
```

Looks like the best model will contain month, noise, site and time as predictors. This is the summary of that model:

```{r model, echo=FALSE}
setwd("C:/Users/bankh/My_Repos/habor-seal/data")
full.data<-read.csv("full.data.csv")
# Load packages
require(MASS) #for glm.nb()

model3<- glm.nb(seals ~ site*noise + month + time, data = full.data)
summary(model3) 
```

* Month and time are significant predictors for how many harbor seals haul-out.

* Site and noise are significant predictors for the number of harbor seals hauled-out. The effect of noise on the number of seals haul-out depends on what site they are located in.

```{r interact}
# What does the interaction between site and noise look like?
interaction.plot(x.factor = full.data$noise, #x-axis variable
                 trace.factor = full.data$site, #variable for lines
                 response = full.data$seals, #y-axis variable
                 fun = median, #metric to plot
                 ylab = "Number of Seals Hauled-out",
                 xlab = "Noise Level (dB)",
                 col = c("pink", "blue"),
                 lty = 1, #line type
                 lwd = 2, #line width
                 trace.label = "Site")
```

