---
title: "Harbor seals analysis"
output: html_document
date: "2022-12-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Waterfront-Marina GLM analysis

In this markdown I will:

1. Find the best distribution to use for GLMs.

2. Run GLMS and AICs to find the most appropriate model and predictors.

3. Create visualization graphs for each site.

### Check Histogram

```{r density, echo=FALSE, message=FALSE}
setwd("C:/Users/bankh/My_Repos/habor-seal/data")
full.data<-read.csv("full.data.csv")
# Load packages
require(ggplot2)

# Check distribution
ggplot(full.data, aes(x=seals)) + 
  geom_histogram(aes(y=..density..), colour="black", fill="white")+
  geom_density()+stat_density(alpha=.2,adjust = 1, fill="#FF6666")+xlab("Number of Seals Hauled-out")+ylab("Density")+theme(panel.background = element_blank())

```

A negative binomial model would fit this data best.

## Create GLMs and find best model with AICc
To test whether noise affects the number of seals hauled-out by site, I will insert an interaction between noise level and site.

```{r glm, echo=FALSE}
setwd("C:/Users/bankh/My_Repos/habor-seal/data")
full.data<-read.csv("full.data.csv")
# Load packages
require(MASS) #for glm.nb()

model1<- glm.nb(seals ~ 1, data = full.data)
model2<- glm.nb(seals ~ site*noise + month + tide + time, data = full.data)
model3<- glm.nb(seals ~ site*noise + month + time, data = full.data)
model4<- glm.nb(seals ~ site*noise + month, data = full.data)
# Create list
models<-list(model1, model2, model3, model4)

## Find the AICc function
source("../code/Functions.R")

## Calculate AICc with glm of models
AICc(models) # Looks like site*noise + month + time are the best predictors
```

Looks like the best model will contain month, noise, site and time as predictors. This is the summary of that model:

```{r model, echo=FALSE}
setwd("C:/Users/bankh/My_Repos/habor-seal/data")
full.data<-read.csv("full.data.csv")
# Load packages
require(MASS) #for glm.nb()

model3<- glm.nb(seals ~ site*noise + month + time, data = full.data)
summary(model3) 
```

Everything is significant.

## Graphs

### Number of Seals Hauled-out vs. Noise Level

#### Waterfront

```{r Wsealsvnoise, echo=FALSE, message=FALSE, warning=FALSE}
# Set working directory here
setwd("C:/Users/bankh/My_Repos/habor-seal/data")
new.w.data<-read.csv("new.w.data.csv")
# Load packages
require(ggplot2)

## Waterfront
ggplot(new.w.data, aes(noise,seals))+geom_point()+geom_smooth(method = "gam", formula =seals~s(noise))+
  stat_smooth(method="gam",colour="black")+xlab("Average Noise Level (dB)")+ylab("Number of Seals Hauled-out")+
  coord_trans(x = "log10")+theme(panel.background = element_blank())
```

#### Marina

```{r Msealsvnoise, echo=FALSE, message=FALSE, warning=FALSE}
# Set working directory here
setwd("C:/Users/bankh/My_Repos/habor-seal/data")
m.data<-read.csv("m.data.csv")
# Load packages
require(ggplot2)

## Marina
ggplot(m.data, aes(noise,seals))+geom_point()+geom_smooth(method = "gam", formula =seals~s(noise))+
  stat_smooth(method="gam",colour="black")+xlab("Average Noise Level (dB)")+ylab("Number of Seals Hauled-out")+
  coord_trans(x = "log10")+theme(panel.background = element_blank())
```

### Number of Seals Hauled-out vs. Month

#### Waterfront

```{r wsealsvmonth, echo=FALSE, message=FALSE}
# Set working directory here
setwd("C:/Users/bankh/My_Repos/habor-seal/data")
new.w.data<-read.csv("new.w.data.csv")
# Load packages
require(ggplot2)

## Waterfront
p<-ggplot(new.w.data, aes(x=month,y =seals,group=month))
p+geom_boxplot(fill="black", alpha=0.2)+
  xlab("Month")+ylab("Number of Seals")+theme(panel.background = element_blank())
```

#### Marina

```{r msealsvmonth, echo=FALSE, message=FALSE}
# Set working directory here
setwd("C:/Users/bankh/My_Repos/habor-seal/data")
m.data<-read.csv("m.data.csv")
# Load packages
require(ggplot2)

## Marina
p<-ggplot(m.data, aes(x=month,y =seals,group=month))
p+geom_boxplot(fill="black", alpha=0.2)+
  xlab("Month")+ylab("Number of Seals")+theme(panel.background = element_blank())
```

