---
title: "Number of Seals Hauled-out Compared to Canidate Model Predictors"
author: "Kyra Bankhead"
output: pdf_document
date: "2022-12-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this markdown I will:

1. Look at how month affects the number of hauled-out seals.

2. Look at how time of day affects the number of hauled-out seals.

## Month vs Seals

#### Waterfront

```{r wsealsvmonth, echo=FALSE, message=FALSE, warning=FALSE}
new.w.data<-read.csv("../data/new.w.data.csv")
# Load packages
require(ggplot2)

## Waterfront
p<-ggplot(new.w.data, aes(x=month,y =seals,group=month))
p+geom_boxplot(fill="black", alpha=0.2)+
  xlab("Month")+ylab("Number of Seals")+theme(panel.background = element_blank())
```

Let's look at the max number of seals and IQR:

```{r wIQR, echo=FALSE, message=FALSE, warning=FALSE}
new.w.data<-read.csv("../data/new.w.data.csv")

summary(new.w.data$seals[new.w.data$month == 7])
IQR(new.w.data$seals[new.w.data$month == 7])

```


#### Marina

```{r msealsvmonth, echo=FALSE, message=FALSE}
m.data<-read.csv("../data/m.data.csv")
# Load packages
require(ggplot2)

## Marina
p<-ggplot(m.data, aes(x=month,y =seals,group=month))
p+geom_boxplot(fill="black", alpha=0.2)+
  xlab("Month")+ylab("Number of Seals")+theme(panel.background = element_blank())
```

Let's look at the max number of seals and IQR:

```{r mIQR, echo=FALSE, message=FALSE, warning=FALSE}
m.data<-read.csv("../data/m.data.csv")

summary(m.data$seals[m.data$month == 8])
IQR(m.data$seals[m.data$month == 8])

```

## Time of Day vs Seals

#### Waterfront

```{r wsealsvtime}
new.w.data<-read.csv("../data/new.w.data.csv")
require(ggplot2)

# Plot 
p<-ggplot(new.w.data, aes(x = time, y = seals, group = time))
p+geom_boxplot(fill="black", alpha=0.2)+
  xlab("Time of Day")+ylab("Number of Seals")+theme(panel.background = element_blank())

```

Let's look at the max number of seals and IQR:

```{r wtimeIQR, echo=FALSE, message=FALSE, warning=FALSE}
new.w.data<-read.csv("../data/new.w.data.csv")

summary(new.w.data$seals[new.w.data$time == 16])
IQR(new.w.data$seals[new.w.data$time == 16])

```

#### Marina

```{r msealsvtime}
m.data<-read.csv("../data/m.data.csv")
require(ggplot2)

# Plot 
p<-ggplot(m.data, aes(x = time, y = seals, group = time))
p+geom_boxplot(fill="black", alpha=0.2)+
  xlab("Time of Day")+ylab("Number of Seals")+theme(panel.background = element_blank())

```

Let's look at the max number of seals and IQR:

```{r mtimeIQR, echo=FALSE, message=FALSE, warning=FALSE}
m.data<-read.csv("../data/m.data.csv")

summary(m.data$seals[m.data$time == 8])
IQR(m.data$seals[m.data$time == 8])

```
