---
title: "Noise Difference Between Sites"
author: "Kyra Bankhead"
output: pdf_document
date: "2022-12-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this markdown I will:

1. Check the assumptions of a t-test.

2. Run a wilcox test on the noise levels between sites.

3. Graph the differences in noise levels between sites.

## Check t-test assumptions

### Are the noise levels normal within sites?

```{r norm, echo=FALSE}
full.data<-read.csv("../data/full.data.csv")

# waterfront
hist(full.data$noise[full.data$site == "waterfront"], breaks = 15, main = "Waterfront", xlab = "Noise Level (dB)")
# marina
hist(full.data$noise[full.data$site == "marina"], breaks = 15, main = "Marina", xlab = "Noise Level (dB)")
```

The marina site has nonnormality in noise levels.


### Are the noise level variences equal between sites?

```{r var}
full.data<-read.csv("../data/full.data.csv")

# waterfront
var(full.data$noise[full.data$site == "waterfront"])

# marina
var(full.data$noise[full.data$site == "marina"])
```

Variences are not equal. Therefore we need to run a wilcox test (U-test).


## Check Wilcox Test

```{r wilcox, echo=FALSE}
full.data<-read.csv("../data/full.data.csv")

#U-test
wilcox.test(noise ~ site, data = full.data,
            exact = FALSE)
res<-wilcox.test(full.data$noise[full.data$site == "waterfront"], full.data$noise[full.data$site == "marina"])
res$p.value
```
The noise levels between sites are significantly different.


## Check average and standard devition noise levels within sites

```{r avg, echo=TRUE}
new.w.data<-read.csv("../data/new.w.data.csv")
m.data<-read.csv("../data/m.data.csv")

# waterfront
mean(new.w.data$noise)
sd(new.w.data$noise)

# marina
mean(m.data$noise)
sd(m.data$noise)
```

## Graph these differences in a violin plot

```{r graph, echo=FALSE, message=FALSE, warning=FALSE}
full.data<-read.csv("../data/full.data.csv")
require(ggplot2)

#Violin of Noise
x.order <- c('Waterfront', 'Marina')
ggplot(full.data, aes(x = site, y = noise)) +
  geom_violin(fill = "grey")+ geom_boxplot(width = .2)+xlab("Site")+ylab("Average Noise Level (dB)")+scale_x_discrete(limit = c("waterfront", "marina"),labels = c("Waterfront","Marina"))+theme(panel.background = element_blank())

```

